---
title: "**Injuries Stratified by Sex**"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r, setup, echo = F, message = FALSE, warning=FALSE}
library(tidyverse)
library(p8105.datasets)
library(hexbin)
library(rvest)
library(ggridges)
library(patchwork)
knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(0817)
```


```{r, include=FALSE}
data("injuries")
data("population")
data("products")

head(population)
head(injuries)
head(products)
```


```{r, echo = F, message = FALSE, warning = FALSE}

test = injuries %>% 
  select(code = prod1, everything())

new = merge(test, products, by = "code")
```



# Distribution of injuries by age stratified by Sex

```{r}
ggplot(new, aes(x = age, fill = sex)) + 
  geom_histogram()
```

The figure above shows the distribution of injury cases by age stratified by gender. By the overlap in distributions, we see that there are more female injury cases compare to male injury cases.

# Plot of most common injured body parts (Overall)

```{r}
injuries %>% 
  count(body_part) %>% 
  	filter(n > 100000) %>% 
  arrange(desc(n)) %>% 
  mutate(
		body_part = factor(body_part),
		body_part = fct_reorder(body_part, n)
	) %>% 
  ggplot(aes(x = body_part, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
    labs(x="Body Part", y="N")
```

## Top 3 causes for the top three injured body parts

```{r}

new %>% 
	filter(body_part %in% c("Head", "Face", "Finger")) %>%
	group_by(body_part) %>% 
	count(title) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(body_part, rank) %>% 
	knitr::kable()

```


# Plot of most common injured body part for Males

```{r}
injuries %>% 
  filter(sex == "Male") %>% 
  count(body_part) %>% 
  arrange(desc(n)) %>% 
  mutate(
		body_part = factor(body_part),
		body_part = fct_reorder(body_part, n)
	) %>% 
  ggplot(aes(x = body_part, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
     labs(x="Body Part", y="N")
```

## Table of top 3 causes for the top three injured body parts for Males

```{r}


new %>% 
	filter(sex == "Male", body_part %in% c("Head", "Finger", "Face")) %>%
	group_by(body_part) %>% 
	count(title) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(body_part, rank) %>% 
	knitr::kable()
```

# Table of top 3 causes for the top three injured body parts for Females

```{r}

# Plot of most common injured body part for Females

injuries %>% 
  filter(sex == "Female") %>% 
  count(body_part) %>% 
  arrange(desc(n)) %>% 
  mutate(
		body_part = factor(body_part),
		body_part = fct_reorder(body_part, n)
	) %>% 
  ggplot(aes(x = body_part, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```


## Table of top 3 causes for the top three injured body parts for Females

```{r}
new %>% 
	filter(sex == "Female", body_part %in% c("Head", "Lower Trunk", "Face")) %>%
	group_by(body_part) %>% 
	count(title) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(body_part, rank) %>% 
	knitr::kable()

```


